---
title: "Air Passenger Dataset"
output: word_document
---



#Loading the dataset

```{r}

data("AirPassengers")
AP.ts<- AirPassengers


```


We can check the Class

```{r}

class(AP.ts)
str(AP.ts)

```




The Maximum value



```{r}
max(AP.ts)


```




The data set contains the AirPassengers travel rate from the year 1949 to 1960

```{r}
start(AP.ts); end(AP.ts); frequency(AP.ts);


```

It starts from year 1949 (January); It ends at 1960 (December) and the Frequency interval is 12


Time series plot:

```{r}

plot(AP.ts)
plot(AP.ts, ylab="Passengers (1000s)", type="o", pch=20)
abline(reg=lm(AirPassengers~time(AirPassengers)))


```



Time period of the dataset

```{r}
cycle(AP.ts) 

```


```{r}
aggregate(AP.ts) 


```


It adds up all of the months for each year.
```{r}
plot(aggregate(AP.ts))


```



Box Plot
```{r}
boxplot(AP.ts~cycle(AP.ts))


```





```{r}

aggregate(AP.ts, FUN=mean)


```

Average frequesncy per month for each year;  average monthly travel increased from 126 in 1949 to 476 in 1960


We can see that July and August have the highest travel volume.

There is growth in the airline passenges numbers, it increases steadily.
There is periodic Fluctuations so seasonality exists.

We can infer that passengers might prefer to travel during summer holiday, again there is a  downfall, and the number again increases during christmas holiday


Decomposing the data 



```{r}

decompose(AP.ts, type= "multiplicative")

AP.decompM <- decompose(AP.ts, type= "multiplicative")
plot(AP.decompM)

```





Autocorrelation graph

```{r}
acf(AP.ts)

```



It is corellated over time.
There is seasonality, we can a high peak at one year, air travel today is much correlated with air travel one year ago

Blue line is the confidence interval where anything outside the boundary has significant relation.



