---
title: "Time Series Forecasting"
output:
  word_document: default
  html_notebook: default
---

AUTHOR: AJANTIKA PAUL

Loading the dataset:


```{r}

data("AirPassengers")
AP.ts<- AirPassengers
```


The data set contains the AirPassengers travel rate from the year 1949 to 1960

Checking the dataset
```{r}

class(AP.ts)
start(AP.ts) 
end(AP.ts) 
frequency(AP.ts) 
summary(AP.ts)



```

It starts from year 1949 (January); It ends at 1960 (December) and the Frequency interval is 12



#Time period of the dataset

```{r}
cycle(AP.ts) 
aggregate(AP.ts) #adds up all of the months for each year
plot(aggregate(AP.ts))

```

It has trend.

#Box Plot
```{r}
boxplot(AP.ts~cycle(AP.ts))

```

```{r}
aggregate(AP.ts, FUN=mean)

```
Average frequesncy per month for each year;  average monthly travel increased from 126 in 1949 to 476 in 1960


July and August have the highest travel volume

There is growth in the airline passenges numbers, it increases steadily.
There is trend and seasonality.
We can infer that passengers might prefer to travel during summer holiday, again there might be a downfall, and the number can again increase during christmas holiday

Time series plot:



```{r}
plot(AP.ts)
plot(AP.ts, ylab="Passengers (1000s)", type="o", pch=20)
abline(reg=lm(AirPassengers~time(AirPassengers)))

```


We can see from in the time series there seems to be seasonal variation in the number of AirPassenger Travel Rate.



Decomposing the data 


```{r}
decompose(AP.ts)
plot(decompose(AP.ts))
```

The plot above shows the original time series (top), the estimated trend component (second from top), the estimated seasonal component (third from top), and the estimated irregular component (bottom). 


Autocorrelation

```{r}
acf(AP.ts)

```

We can see that there is lot of persistence from one period to another.

It is corellated over time.
There is seasonality, we can a high peak at one year, air travel today is much correlated with air travel one year ago

Blue line is the confidence interval where anything outside the boundary has significant relation.



Holt Winters Method

```{r}
plot(HoltWinters(AP.ts, alpha=0.001, beta=1, gamma=0))



```

Black graph is the original data and red graph is the Holt-Winters
It Fits well.




```{r}
AR.hw <- HoltWinters(AP.ts, seasonal="additive")

summary(AR.hw)



```



Predictions

We are predicting ahead by 10 years

```{r}
AR.predict<-predict(AR.hw, n.ahead=10*12)
ts.plot(AP.ts, AR.predict, lty=1:2)
```


Predicted values of  AirTravel Rates.


```{r}
AR.predict
```




ETS- Error Trend and Seasonality.



```{r}

library(forecast)


ets_forecast <- ets(AP.ts)
ets_forecast

```

We can see that alpha = 0.73 which is close to one which means that more weight is place on the most recent observations. 

ETS(A,Ad,N) - Error type-"A"=Multiplicative,, Tread-additive, Seasonality- Multiplicative

```{r}
plot(ets_forecast)
attributes(ets)
attributes(ets_forecast)
ets_forecast$mse #MSE gives an overall idea of the error occurred during forecasting. 


```



Forecast

To make a forecast of Air Passengers Travel for the year 1961 to 1970 using forecast



```{r}
forecast.ets(ets_forecast, h=120)

```


The forecast ets() function gives  the forecast for a year, a 80% prediction interval for the forecast, and a 95% prediction interval for the forecast. For example, the forecasted Air pasenger travel rate for the year 1970 is about 504,with a 95% prediction interval of (1031).



```{r}
forecast_ets <- forecast.ets(ets_forecast, h=96)
plot(forecast_ets)
```


Here the forecasts are plotted as a blue line, the 80% prediction interval as an dark grey shaded area, and the 95% prediction interval as a light grey shaded area.


Partial autocorrelation
```{r}
pacf(AP.ts)


```



Regression on time dimension


```{r}

AP.reg<-lm(AP.ts~time(AP.ts))
summary(AP.reg)
confint(AP.reg)
acf(resid(AP.reg))
pacf(resid(AP.reg))
```



ARIMA (Autoregressive integrated moving average)
MA(3)


```{r}
AP.ma <- arima(AP.ts, order=c(0,0,3)) #estimates moving average of three time periods of "white noise"

```




```{r}
ar=auto.arima(AP.ts)
summary(ar)
```

ARMA(1,1)

```{r}
AP.arma <- arima(AP.ts, order=c(1,0,1)) #1 prior period of AP, 1 prior period of white noise

```

ARIMA(2,1,2)
This model is composed of 2 prior periods of AP, 2 prior periods of white noise and a first-order difference

```{r}
AP.arima <- arima(AP.ts, order=c(2,1,2)) 
AP.predict<-predict(AP.arima, n.ahead=60)
ts.plot(AP.ts,AP.predict$pred, lty=1:2)


```


#GARCH
#Generalized autoregressive conditioned heteroskedastic


```{r}
library(tseries)
garch(AP.ts, grad="numerical")

```



